{\sf NodIO}, a framework for evolutionary algorithms in JavaScript: first results

### 2016-01-05 special-issue-ci/volunteer-clouds-ieee-ci-si-tr.tex
This paper is going to need a bit of restructuring
Changing the title again. Please @mariosky @pacastillo @unintendedbear take a look. Anyone, in fact. Refs #82 #83

I have put a title in every set of paragraphs to see how we can structure it

\title{Revisiting evolutionary algorithms in JavaScript}

@@

JavaScript is an interpreted language mainly known for its inclusion
as an embedded language for programming internet browsers. This has
inspired its use, for a long time, as a tool for evolutionary
algorithms; several libraries have been published so far and are in
use. However, the last years have seen a resurgence of interest in the
language, becoming one of the most popular and thus spawning the
improvement of its implementations, which form now the base of many
new client-server applications. That is why in this paper we propose a series of
measurements to establish the speed of JavaScript in evolutionary
algorithms that can serve as a baseline for comparison with other
distributed computing experiments. These experiments use different
integer and floating point problems, and prove that the speed of
JavaScript is actually competitive with other languages commonly used
by the evolutionary algorithm practitioner.

@@

The rest of the paper is organized as follows:

%---------------------------------------------------------------
\section{State of the art}
\label{sec:soa}

%---------------------------------------------------------------
\section{Description of the framework}
\label{sec:description}

In general, a distributed volunteer-based evolutionary computation
framework based on the browser is simply a client-server system
whose client is, or it can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself, which % (Paloma) I recommend the use of "which" here
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. The results of this work show that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact,
% faster, but the fact that single-user performance is higher (than
% what?) is over the offset. This is because the number of spontaneous
% volunteers we can gather at whim, including people using their
% mobile phones or tablets, is several orders of magnitude higher,
% which is the objective in this kind of systems." (?) 
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and then compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

@@

In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
cloud or bare metal based volunteer evolutionary computing system
derived from % (Paloma) In order to not sounding too repetitive, could we use "derived from", for example, here?
 the {\sf NodEO} library, whose architecture has been
developed using JavaScript on the client as well as the server.
All parts of the framework are free and available with a free license
from \url{https://github.com/JJ/splash-volunteer}.

@@

The figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the user interface elements of the framework.
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{system.pdf}
\caption{Description of the proposed framework. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server.}
\label{fig:system}
\end{figure}

>>>>

\title{{\sf NodIO}, a framework for evolutionary algorithms in JavaScript: first results}

@@

JavaScript is an interpreted language mainly known for its inclusion
as an embedded language for programming internet browsers. This has
inspired its use, for a long time, as a tool for evolutionary
algorithms, mainly so in browser-based volunteer computing
environments. Several libraries have also been published so far and are in
use. However, the last years have seen a resurgence of interest in the
language, becoming one of the most popular and thus spawning the
improvement of its implementations, which form now the base of many
new client-server applications. That is why in this paper we propose a series of
measurements to establish the speed of JavaScript in evolutionary
algorithms that can serve as a baseline for comparison with other
distributed computing experiments. These experiments use different
integer and floating point problems, and prove that the speed of
JavaScript is actually competitive with other languages commonly used
by the evolutionary algorithm practitioner.

@@

In general, a distributed volunteer-based evolutionary computation
framework based on the browser is simply a client-server system
whose client is, or can be, embedded in the browser via
JavaScript. Since JavaScript is  the only language that is present
across all browsers, the choice was quite clear. Even so, there might
be some issues with the performance of the language itself, which 
 is
why we have made a comparison between JavaScript and other languages
\cite{2015arXiv151101088M}. That paper shows that JavaScript can be
faster than compiled languages such as Scala and, in any case, it provides
a performance that is comparable to other languages usually employed % (Paloma) Adverb
in evolutionary computation and numerical optimization such as Python. 
% (Paloma) Bit of a too long sentence here. Why not changing the ; for a . ?
Compiled languages such as
Java or C might be faster when % (Paloma) Too many "fact" words. There was another one in the previous sentence.
considering the performance of a single-user, 
but this is more than offset by the computational power of
the spontaneous volunteers we can gather at whim, including people
using their mobile phones or tablets. Together, the performance is several orders of magnitude
higher, which is the objective in this kind of systems.
% (Paloma) This sentence needs re-writing
% "Other compiled languages such as Java or C might be, in fact,
% faster, but the fact that single-user performance is higher (than
% what?) is over the offset. This is because the number of spontaneous
% volunteers we can gather at whim, including people using their
% mobile phones or tablets, is several orders of magnitude higher,
% which is the objective in this kind of systems." (?) 
 JavaScript is
quite a popular language nowadays, providing compilers from a number
of languages, many of them in their family (like CoffeeScript), but
many other outside, so researchers can, in fact, write their fitness
function in Python, Erlang, Scala, or even Java, and then compile it to
JavaScript. For completeness, an experiment comparing the performance of
a JavaScript implementation versus Java and Matlab is included in section
\ref{sec:w2}.

%%% Issues with volunteer based systems

@@

%%% Versions of the framework

In this sense, in this paper we propose several versions of the {\sf NodIO} framework, a
cloud or bare metal based volunteer evolutionary computing system
derived from the {\sf NodEO} library, whose architecture has been
developed using JavaScript on the client as well as the server.
All parts of the framework are free and available with a free license
from \url{https://github.com/JJ/splash-volunteer}.

@@

% Description of the framework
The figure \ref{fig:system} describes the general system architecture and
algorithm behavior. Different web technologies, such as JQuery or {\tt
  Chart.js} have
been used to build the user interface elements of the framework.
\begin{figure}[!t]
\centering
\includegraphics[width=4in]{system.pdf}
\caption{Description of the proposed framework. Clients execute a JavaScript EA
  in the browser, which, every 100 generations, sends the best
  individual and receives a random one from the server.}
\label{fig:system}
\end{figure}
